% 
% Permission is granted to copy, distribute and/or modify this document
% under the terms of the GNU Free Documentation License, Version 1.2
% or any later version published by the Free Software Foundation;
% with no Invariant Sections, no Front-Cover Texts, and no Back-Cover
% Texts.  A copy of the license is included in the section entitled "GNU
% Free Documentation License".

\section{Validation}

This section provides some insight on the performance of the present implementation of the ADS-2 method (and its variants) on simple academical benchmarks. The method will be compared to crude MC. The benchmarks expose the total number of calls to the model $N$, the failure probability $P_F$, the generalized reliability index $\beta$ and the coefficient of variation $CV$.\par

\subsection{Flood problem}

The purpose of this example is to estimate the risk of a flood to run over a dyke.\par

\subsubsection{Problem statement}

The limit-state function is defined as follows:
\begin{eqnarray*}
  L = 5000.\\
  B = 300.\\
  H_d = 0.\\
  Z_b = 55.5\\
  S(Q, K_s, Z_v, Z_m) = Z_v + \left( \frac{Q}{K_s B \sqrt{\frac{Z_m - Z_v}{L}}} \right) ^\frac{3}{5} - (H_d + Z_b)\\
  S < 0
\end{eqnarray*}
The random variables distribution are given in Table~\ref{tab:flood_prob}. All components are assumed independent.\par

\begin{table}
    \centering
    \begin{tabular}{lrr}
        \hline
        Variable & Distribution & \text{Parameters} \\
        \hline
        Q & Gumbel   & $\alpha = \frac{1}{558}, \beta=1013$  \\
        $K_s$ & Normal   & $\mu=30, \sigma= 7.5$ \\
        $Z_v$ & Triangular   & $a=49, m=50, b=51$ \\
        $Z_m$ & Triangular   & $a=54, m=55, b=56$ \\
        \hline
    \end{tabular}
    \caption{Probabilistic model for the flood problem.}
    \label{tab:flood_prob}
\end{table}

\newpage

\subsubsection{Results}

Let use the method for several dyke heights $H_d$, with $n=10^5$ Monte Carlo simulations (MCS) and thumb rule $n_{ADS}=10^5/20$ for ADS-2 which is expected to yield $n < 10^5$ evaluations of the limit-state function (20 per sampled direction). The computational budget is split as follows: $\gamma_1,\,\gamma_2 = 0.5$.\par

\begin{table}
    \centering
    \begin{tabular}{ccccc}
        \hline
        Method & $H_d$ & $n$ & $\hat P_f$ & $\hat {CV}$ \\
        \hline
        MCS & 0. & $10^5$ & $1.18 \times 10^{-2}$ & $2.89 \times 10^{-2}$\\
        MCS & 2. & $10^5$ & $1.30 \times 10^{-3}$ & $8.76 \times 10^{-2}$\\
        MCS & 5. & $10^5$ & $2.00 \times 10^{-4}$ & $2.24 \times 10^{-1}$\\
        MCS & 8. & $10^5$ & $9.00 \times 10^{-5}$ & $3.33 \times 10^{-1}$\\
        ADS-2 & 0.& $66,395$ & $1.30 \times 10^{-2}$ & $2.56 \times 10^{-3}$\\
        ADS-2 & 2.& $83,573$ & $1.22 \times 10^{-3}$ & $2.40 \times 10^{-3}$\\
        ADS-2 & 5.& $99,781$ & $1.92 \times 10^{-4}$ & $3.38 \times 10^{-3}$\\
        ADS-2 & 8.& $94,993$ & $6.65 \times 10^{-5}$ & $3.53 \times 10^{-3}$\\
        \hline
    \end{tabular}
    \caption{Benchmark results for the flood model.}
\end{table}

The ADS-2 method gives much better coefficients of variation than the standard Monte Carlo simulation for approximately the same number of calls to the limit-state function.\par

\newpage

\subsection{The hyperplane problem}

This parametric example allows to spread the failure domain over either one or several quadrants. This feature enables an emphasis of the performance of the dimension reduction (ADS-2+) and design point (DP-ADS-2) variants.\par

\subsubsection{Problem statement}

The hyperplane limit-state function is defined as follows:
\begin{eqnarray*}
  H(U_1, \dots, U_d) = \sum\limits_{i=1}^d \alpha_i U_i\\
  H < k\\
\end{eqnarray*}
where $U_i, i = 1, \ldots, d$ are independent, zero-mean and unit-variance normal random variables.\par

The failure probability comes analytically in this case and reads:
\begin{equation*}
    P_f = \Phi\left(\frac{k}{\sqrt{\sum\limits_{i=1}^d \alpha_i^2}}\right).
\end{equation*}
Hence given $\alpha_i$'s values, we can choose $k$ for a given target probability $P_f$:
\begin{equation*}
    k = \Phi^{-1}(P_f)\,\sqrt{\sum\limits_{i=1}^d \alpha_i^2}.
\end{equation*}

\subsubsection{Results}

For this case we will use $n_{ADS}=\frac{10^5}{11}$ as above. A first hyperplane example (H1) is defined where $\alpha_i = 1$ with $n = 10^5$ for several failure levels $P_f = 10^{-3}, 10^{-4}, 10^{-5}$ in dimension $d = 5$. The total computational budget is split as follows: $\gamma_1,\,\gamma_2 = 0.5$. The results are provided in Table~\ref{tab:H1}.\par

\begin{table}
    \centering
    \begin{tabular}{cccccc}
        \hline
        Method & $P_f$ & $k$ & n & $\hat P_f$ & $\hat {CV}$ \\
        \hline
        MC & $1 \times 10^{-3}$ & $-15.45$ & $10^5$ & $9.00 \times 10^{-4}$ & $1.05 \times 10^{-1}$ \\
        MC & $1 \times 10^{-4}$ & $-18.60$ & $10^5$ & $8.00 \times 10^{-5}$ & $3.54 \times 10^{-1}$\\
        MC & $1 \times 10^{-5}$ & $-21.32$ & $10^5$ & $1.00 \times 10^{-5}$ & $1.00 \times 10^{0}$\\
        ADS-2 & $1 \times 10^{-3}$ & $-15.45$ & $99,671$ & $2.20 \times 10^{-3}$ & $4.66 \times 10^{-4}$\\
        ADS-2 & $1 \times 10^{-4}$ & $-18.60$ & $99,301$ & $1.06 \times 10^{-4}$ & $1.13 \times 10^{-3}$\\
        ADS-2 & $1 \times 10^{-5}$ & $-21.32$ & $10,0943$ & $9.91 \times 10^{-6}$ & $1.63 \times 10^{-3}$\\
        \hline
    \end{tabular}
    \label{tab:H1}
    \caption{Benchmark results for the first hyperplane problem (H1).}
\end{table}

The ADS-2 method gives much better coefficients of variation than the standard Monte Carlo simulation for approximately the same number of calls to the limit-state function.\par

A second hyperplane example (H3) where $\alpha_1, \dots, \alpha_q = 1, \alpha_{q+1}, \dots, \alpha_{d} = 0$ with $q = 2$, $n = 10^5$, $n_{ADS} = \frac{10^5}{16}$, $P_f = 1 \times 10^{-4}$ (hence $k \approx -18.60$) in dimension $d=5$, $\gamma_1,\gamma_2=0.5$. The ADS-2+ variant will use $d' = 2 < 5$ as the number of stratified dimensions, and the DP-ADS-2 variant will use the $\alpha_i$ coefficients as the quadrant orientation. The results are provided in Table~\ref{tab:H3}.\par

\begin{table}
    \centering
    \begin{tabular}{cccc}
        \hline
        Method & n & $\hat P_f$ & $\hat {CV}$ \\
        \hline
        MC & $10^5$ & $5.00 \times 10^{-5}$ & $4.47 \times 10^{-1}$ \\
        ADS-2 & $104,555$ & $1.04 \times 10^{-4}$ & $3.34 \times 10^{-3}$ \\
        ADS-2+ & $93,316$ & $9.48 \times 10^{-5}$ & $1.79 \times 10^{-3}$ \\
        DPADS-2 & $98,147$ & $9.92 \times 10^{-5}$ & $2.93 \times 10^{-3}$ \\
        \hline
    \end{tabular}
    \caption{Benchmark results for the second hyperplane problem (H3).}
    \label{tab:H3}
\end{table}

Both ADS-2+ and DP-ADS-2 variants yield slight improvements in terms of number of calls to the limit state function comparing to the vanilla ADS-2 method.\par
